---
title: "Advanced sl3 Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced sl3 Integration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The high-level API covers common analyses, but the original advanced `hte3`
interfaces remain available for fine-grained control over nuisance learners
or meta-learners.

Use the advanced API to:

- construct the `hte3_Task` manually
- specify the exact low-level `Lrnr_*` candidates
- call `cross_validate_cate()` or `cross_validate_crr()` yourself

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 100, seed = 4)

task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)

learner <- Lrnr_cate_DR$new(base_learner = Lrnr_mean$new())
trained <- learner$train(task)

head(predict_hte3(trained, data))
```

## Low-Level CATE Portfolios

You can cross-validate across low-level CATE learners directly:

```{r}
cate_cv <- cross_validate_cate(
  list(
    Lrnr_cate_DR$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_R$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
  ),
  task,
  cv_control = list(V = 2)
)

cate_cv$coefficients
head(cate_cv$lrnr_sl$predict(task))
```

## Low-Level CRR Portfolios

The analogous CRR workflow is:

```{r}
crr_task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)

crr_cv <- cross_validate_crr(
  list(
    Lrnr_crr_IPW$new(base_learner = Lrnr_mean$new()),
    Lrnr_crr_T$new(base_learner = Lrnr_mean$new()),
    Lrnr_crr_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
  ),
  crr_task,
  cv_control = list(V = 2)
)

crr_cv$coefficients
head(crr_cv$lrnr_sl$predict(crr_task))
```

## How Advanced CV Fits Together

- `cross_fit_and_cv` in `make_hte3_Task_tx()` controls nuisance estimation.
- `cross_validate_cate()` and `cross_validate_crr()` control selection among
  HTE learners.
- If a candidate learner itself contains a stacked `sl3` base learner, that
  base learner can internally cross-validate as well.

The advanced API is lower-level and closer to the original research
implementation. Use it when full control over candidate portfolios is
required.

For continuous-treatment CATE, the low-level `Lrnr_cate_R` path is the one
implemented in the package. It follows the partially linear effect-model view
with an `A * tau(X)` term.
