---
title: "Advanced sl3 Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced sl3 Integration}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Read this article if you want the low-level API: explicit task construction,
manual learner choice, or direct calls to `cross_validate_cate()` and
`cross_validate_crr()`.

## Build the Task Explicitly

Use `make_hte3_Task_tx()` when you want to supply the nuisance learners
yourself or work directly with the lower-level learner objects.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 100, seed = 4)

task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)
```

## Train a Learner Directly

```{r, warning = FALSE}
learner <- Lrnr_cate_DR$new(base_learner = Lrnr_mean$new())
trained <- learner$train(task)

head(predict_hte3(trained, data))
```

This is the right level when you want the exact learner object rather than the
wrapper returned by `fit_cate()` or `fit_crr()`.

## Low-Level Cross-Validation

You can cross-validate low-level learner portfolios directly.

```{r, warning = FALSE}
cate_cv <- cross_validate_cate(
  list(
    Lrnr_cate_DR$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_R$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
  ),
  task,
  cv_control = list(V = 2)
)

cate_cv$coefficients
head(cate_cv$lrnr_sl$predict(task))
```

The analogous CRR workflow is:

```{r}
crr_task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)

crr_cv <- suppressWarnings(
  cross_validate_crr(
    list(
      Lrnr_crr_IPW$new(base_learner = Lrnr_mean$new()),
      Lrnr_crr_T$new(base_learner = Lrnr_mean$new()),
      Lrnr_crr_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
    ),
    crr_task,
    cv_control = list(V = 2)
  )
)

crr_cv$coefficients
head(crr_cv$lrnr_sl$predict(crr_task))
```

## Important Constraints

- `cross_fit_and_cv` in `make_hte3_Task_tx()` controls nuisance estimation at
  the task stage.
- If `modifiers = V` and `confounders = W` with `V` a strict subset of `W`,
  the reduced-modifier target is `E[Y(1) - Y(0) | V] = E[tau(W) | V]`.
- In that reduced-modifier setting, the current low-level R-learner targets an
  overlap-weighted projection rather than the unweighted `V`-conditional CATE.
- For continuous-treatment CATE, the low-level `Lrnr_cate_R` path is the one
  currently implemented.

For EP-specific sieve construction and tuning, see
[EP-Learner Sieve Tuning](ep-learner.html). If you are building a new learner,
see [Extending hte3 with Custom Learners](extending-hte3.html).
