---
title: "CATE Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CATE Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Read this article if you are fitting conditional average treatment effects
with `hte_task()` and `fit_cate()` and need to choose among the supported CATE
learner families.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 150, seed = 2)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

dr_model <- fit_cate(
  task,
  method = "dr",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(dr_model, data))
summary(dr_model)
```

## Choose the Wrapper Method

`fit_cate()` supports `method = c("dr", "r", "t", "ep")`.

- `method = "dr"` is the main difference-scale default when treatment is
  binary or categorical and you want a target-aligned CATE fit that uses both
  propensity and outcome nuisance estimates.
- `method = "r"` is the residual-on-residual path and the current
  continuous-treatment option. In reduced-modifier settings, it targets an
  overlap-weighted projection rather than the unweighted
  `E[Y(1) - Y(0) | V]`.
- `method = "t"` is the baseline that fits outcome models by treatment arm and
  subtracts them.
- `method = "ep"` is the sieve-based plug-in path. Use it when you want the
  EP family; see [EP-Learner Sieve Tuning](ep-learner.html) for basis-size
  tuning and EP-DR versus EP-R details.

## Reduced-Modifier Targets

When `modifiers = V` and `confounders = W` with `V` a strict subset of `W`,
the target of interest is `E[Y(1) - Y(0) | V] = E[tau(W) | V]`.

In the supported binary/categorical-treatment setting:

- DR, EP, and the default two-stage T-learner align with that target
- the current R-learner instead targets an overlap-weighted projection onto
  functions of `V`

This is why `fit_cate(..., method = "r")` warns in reduced-modifier settings.

## Compare CATE Learner Families

Wrapper-level CATE cross-validation is controlled by `cross_validate` and
`cv_control`.

```{r, warning = FALSE}
cv_model <- fit_cate(
  task,
  method = c("dr", "r", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

cv_model
head(predict(cv_model, data))
```

`summary(cv_model)` reports the selected learner. If the selected learner is
an EP candidate, the summary also reports the selected basis size and any EP
selection metadata.

## Important Constraints

- Continuous-treatment CATE currently supports `method = "r"` only, using the
  partially linear effect model `A * tau(X)`.
- Use `sieve_num_basis` when you want one fixed EP basis size and
  `sieve_basis_grid` when you want wrapper-level EP basis-size selection.
- For direct control over candidate learner objects, see
  [Advanced sl3 Integration](advanced-sl3.html).
