---
title: "CATE Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CATE Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette shows the high-level workflow for conditional average treatment
effect estimation.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 150, seed = 2)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

dr_model <- fit_cate(
  task,
  method = "dr",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

t_model <- fit_cate(
  task,
  method = "t",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(dr_model, data))
head(predict(t_model, data))
```

The basic pattern is:

- build a task with `hte_task()`
- choose a CATE learner family with `method`
- optionally use `cross_validate = TRUE` to select among candidates

## Cross-Validation Across Learner Families

Wrapper-level CATE cross-validation is controlled by `cross_validate` and
`cv_control`.

```{r}
cv_model <- fit_cate(
  task,
  method = c("dr", "r", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

cv_model
head(predict(cv_model, data))
```

This portfolio example evaluates:

- the DR learner
- the R learner
- a portfolio of EP learners with different sieve basis sizes

The outer selection loss is handled by `cross_validate_cate()` using the
DR-style CATE selector.

## Low-Level Cross-Validation

For explicit control over candidate learner objects, use
`cross_validate_cate()` directly.

```{r}
cv_fit <- cross_validate_cate(
  list(
    Lrnr_cate_DR$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_R$new(base_learner = Lrnr_mean$new()),
    Lrnr_cate_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
  ),
  task,
  cv_control = list(V = 2)
)

cv_fit$coefficients
head(cv_fit$lrnr_sl$predict(task))
```

## Notes

- `cross_fit` in `hte_task()` controls nuisance estimation and is separate
  from learner-level cross-validation.
- Continuous-treatment CATE tasks currently support `method = "r"` only, via
  the partially linear R-learner effect model `A * tau(X)`.
- For EP-learner tuning and more direct `Lrnr_*` control, see the advanced
  vignette.
