---
title: "CRR Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CRR Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Read this article if you are fitting conditional risk ratios with
`hte_task()` and `fit_crr()` and need to choose among the supported CRR
learner families.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 150, seed = 3)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

crr_model <- fit_crr(
  task,
  method = "ipw",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(crr_model, data))
summary(crr_model)
```

## Choose the Wrapper Method

`fit_crr()` supports `method = c("ep", "ipw", "t")`.

- `method = "ep"` is the default flexible CRR path.
- `method = "ipw"` is the weighting-based benchmark and places more emphasis
  on the propensity model.
- `method = "t"` is the baseline built from arm-specific outcome models on the
  risk-ratio scale.

The wrapper pattern is the same as for CATE: build an `hte3_Task`, choose the
CRR learner family with `method`, and optionally compare candidates with
`cross_validate = TRUE`.

## Compare CRR Learner Families

```{r, warning = FALSE}
cv_model <- fit_crr(
  task,
  method = c("ipw", "t", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

cv_model
head(predict(cv_model, data))
```

If `method = "ep"` and `cross_validate = TRUE`, the wrapper also compares EP
fits across basis sizes. Use `sieve_basis_grid` when you want to control that
grid explicitly, and use `sieve_num_basis` for one fixed EP fit.

## Important Constraints

- CRR workflows require a non-negative outcome.
- `cross_fit` in `hte_task()` controls nuisance estimation and is separate
  from wrapper-level learner selection.
- For low-level CRR learner portfolios, see
  [Advanced sl3 Integration](advanced-sl3.html).
