---
title: "CRR Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CRR Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette shows the production workflow for conditional risk-ratio style
targets using a non-negative outcome.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 150, seed = 3)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

crr_model <- fit_crr(
  task,
  method = "ipw",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(crr_model, data))
summary(crr_model)
```

The production pattern is the same as for CATE:

- build an `hte3_Task` with `hte_task()`
- choose a CRR learner family with `method`
- optionally select among candidate learners with `cross_validate = TRUE`

## Cross-Validation Across Learner Families

```{r}
cv_model <- fit_crr(
  task,
  method = c("ipw", "t", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

cv_model
head(predict(cv_model, data))
```

This portfolio evaluates:

- the IPW CRR learner
- the T learner on the log risk-ratio scale
- a portfolio of EP learners with different sieve basis sizes

The outer selection loss is handled by `cross_validate_crr()` using the CRR
selector.

## Low-Level Cross-Validation

If you want to control the exact learner list directly, use
`cross_validate_crr()`.

```{r}
cv_fit <- cross_validate_crr(
  list(
    Lrnr_crr_IPW$new(base_learner = Lrnr_mean$new()),
    Lrnr_crr_T$new(base_learner = Lrnr_mean$new()),
    Lrnr_crr_EP$new(base_learner = Lrnr_mean$new(), sieve_num_basis = 4)
  ),
  task,
  cv_control = list(V = 2)
)

cv_fit$coefficients
head(cv_fit$lrnr_sl$predict(task))
```

## Notes

- CRR workflows require a non-negative outcome.
- `cross_fit` in `hte_task()` controls nuisance estimation and is separate
  from HTE learner selection.
- If `method = "ep"` and `cross_validate = TRUE`, the wrapper performs
  basis-size selection across EP candidates automatically.
