---
title: "EP-Learner Sieve Tuning"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EP-Learner Sieve Tuning}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette focuses on the EP-learner. It explains what the `Sieve`
package is doing inside `hte3`, how the sieve basis is constructed and
ordered, and how to reproduce the paper-style sieve cross-validation grids.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 120, seed = 4)

task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)

crr_task <- make_hte3_Task_tx(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  learner_pi = Lrnr_mean$new(),
  learner_mu = Lrnr_mean$new(),
  learner_m = Lrnr_mean$new(),
  cross_fit_and_cv = FALSE
)
```

## What `Sieve` Does in `hte3`

EP-learner starts from arm-specific outcome regressions and then debiases
them over a sieve basis defined on the modifier set `V`. In `hte3`, the
`Sieve` package is used only to build that basis matrix. The EP update
itself is then fit by `glmnet` inside `Lrnr_cate_EP` and `Lrnr_crr_EP`.

Before evaluating the basis, `Sieve` rescales each modifier to `[0, 1]`. In
the current `hte3` implementation, the sieve family is fixed to cosine basis
functions, so the univariate building blocks are `1, cos(pi x), cos(2 pi x),
...`. Multivariate basis functions are tensor products of those univariate
terms across modifiers.

The package does not currently expose alternative `Sieve` basis families at
the wrapper or learner level. The practical tuning knobs are:

- `sieve_num_basis`: sieve size
- `sieve_interaction_order`: maximum interaction order
- `screen_basis_with_lasso`: experimental CATE-only alternative to explicit
  sieve-grid cross-validation

If `sieve_num_basis = NULL`, `hte3` uses the default
`ceiling(n^(1/3) * d)`, where `n` is the sample size and `d` is the number
of modifiers.

## How the Basis Is Ordered

`Sieve` orders the multivariate basis by increasing index-product
complexity. In practice, this means the constant term comes first, then
lower-frequency univariate terms, then higher-frequency and interaction
terms.

The argument `sieve_interaction_order` limits how many coordinates can vary
inside a single tensor-product basis term:

- `sieve_interaction_order = 1` gives an additive sieve
- `sieve_interaction_order = 2` allows pairwise interactions
- `sieve_interaction_order = 3` allows up to three-way interactions

So with two modifiers, the early cosine basis terms look like the constant,
then terms like `cos(pi x1)` and `cos(pi x2)`, then higher-frequency terms,
then interaction terms such as `cos(pi x1) * cos(pi x2)`.

## Fitting a Single EP-Learner

This is the most direct low-level EP fit:

```{r}
ep_fit <- Lrnr_cate_EP$new(
  base_learner = Lrnr_mean$new(),
  sieve_num_basis = 4,
  sieve_interaction_order = 2
)$train(task)

head(predict_hte3(ep_fit, data))
```

In practice, the paper often paired EP-learner with richer base-learner
stacks. The examples below keep `Lrnr_mean$new()` so the sieve tuning pattern
is easy to read; you can replace `base_learner` with any `sl3` learner or
stack.

## Paper-Faithful CATE Sieve CV

The main low-dimensional CATE simulations in the paper used a sieve grid of
`3d, 4d, ..., 8d` with `sieve_interaction_order = 3`.

```{r}
d <- length(task$npsem$modifiers$variables)
paper_cate_basis_grid <- (3:8) * d

paper_cate_ep_stack <- lapply(paper_cate_basis_grid, function(basis_size) {
  Lrnr_cate_EP$new(
    base_learner = Lrnr_mean$new(),
    sieve_num_basis = basis_size,
    sieve_interaction_order = 3
  )
})

paper_cate_cv <- cross_validate_cate(
  paper_cate_ep_stack,
  task,
  cv_control = list(V = 5)
)

paper_cate_basis_grid
paper_cate_cv$coefficients
```

The high-level wrapper `fit_cate(..., method = "ep", cross_validate = TRUE)`
uses a convenience grid. If you want paper-faithful tuning, use an explicit
EP stack like the one above.

For the high-dimensional CATE simulations, the paper used a more
conservative additive sieve:

```{r, eval = FALSE}
d <- length(task$npsem$modifiers$variables)
paper_cate_highdim_grid <- (1:6) * d

paper_cate_highdim_ep_stack <- lapply(
  paper_cate_highdim_grid,
  function(basis_size) {
    Lrnr_cate_EP$new(
      base_learner = Lrnr_mean$new(),
      sieve_num_basis = basis_size,
      sieve_interaction_order = 1
    )
  }
)

paper_cate_highdim_cv <- cross_validate_cate(
  paper_cate_highdim_ep_stack,
  task,
  cv_control = list(V = 5)
)
```

## Paper-Faithful CRR Sieve CV

The main CRR simulations used the same low-dimensional sieve grid
`3d, 4d, ..., 8d` with `sieve_interaction_order = 3`.

```{r}
d_crr <- length(crr_task$npsem$modifiers$variables)
paper_crr_basis_grid <- (3:8) * d_crr

paper_crr_ep_stack <- lapply(paper_crr_basis_grid, function(basis_size) {
  Lrnr_crr_EP$new(
    base_learner = Lrnr_mean$new(),
    sieve_num_basis = basis_size,
    sieve_interaction_order = 3
  )
})

paper_crr_cv <- cross_validate_crr(
  paper_crr_ep_stack,
  crr_task,
  cv_control = list(V = 5)
)

paper_crr_basis_grid
paper_crr_cv$coefficients
```

## Notes

- The paper-faithful path is explicit sieve-grid selection with
  `cross_validate_cate()` or `cross_validate_crr()`.
- For CATE, `screen_basis_with_lasso = TRUE` is an experimental alternative
  when explicit sieve CV is too expensive.
- For broader low-level `sl3` portfolio control beyond EP-learner, see the
  `advanced-sl3` vignette.
