---
title: "Causal Forests with GRF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal Forests with GRF}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Read this article if you want the optional `grf` wrappers rather than the
standard `hte_task()` plus `fit_*()` workflow. The GRF path keeps the same
`hte3_model` output interface, but it uses `grf` as the forest engine and
accepts raw data plus optional nuisance estimates directly.

```{r setup}
library(hte3)
```

## When to Use This Workflow

Use the GRF wrappers when all of the following are true:

- you want a forest-based workflow built around the optional `grf` package
- you want to call a wrapper directly on raw data instead of building
  `hte_task()` first
- you still want an `hte3_model` that works with `predict()` and `summary()`

## Choose the Wrapper from the Estimand

- Use `grf_cate()` when the target is a conditional mean difference.
- Use `grf_crr()` when the target is a conditional risk ratio and the outcome
  is non-negative.

## Choose the CATE Method

`grf_cate()` supports `method = c("dr", "r", "ep")`.

- `method = "r"` is the direct causal-forest path and trains
  `grf::causal_forest()`.
- `method = "dr"` uses the DR-learner with GRF nuisance and base learners.
- `method = "ep"` uses the EP-learner with GRF nuisance and base learners.

Within the EP family, `ep_targeting_style = "dr"` selects EP-DR and
`ep_targeting_style = "r"` selects EP-R. If you pass
`ep_targeting_style = c("dr", "r")`, wrapper-level cross-validation can
compare both EP variants.

If you omit both `method` and `cross_validate`, `grf_cate()` compares the
default EP-R candidate and the direct GRF R-learner.

```{r cate-direct-r, eval = FALSE}
data <- hte3_example_data(n = 200, seed = 1)

fit <- grf_cate(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  method = "r",
  cross_fit = FALSE
)

head(predict(fit, data))
summary(fit)
```

## `grf_crr()`

`grf_crr()` supports `method = c("ep", "ipw", "t")`. If you omit `method`,
the wrapper fits one EP model.

```{r crr-example, eval = FALSE}
fit <- grf_crr(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  method = "ep",
  cross_fit = FALSE
)

head(predict(fit, data))
summary(fit)
```

## Direct Nuisance Inputs

The GRF wrappers accept direct nuisance estimates:

- `mu.hat`: an `n x 2` matrix ordered as `(control, treatment)`
- `pi.hat`: a length-`n` vector for `P(A = treatment_level | X)`
- `m.hat`: an optional length-`n` vector of marginal outcome means

If `m.hat` is omitted and both `mu.hat` and `pi.hat` are supplied, `hte3`
computes `m.hat` internally.

```{r direct-nuisances, eval = FALSE}
mu.hat <- cbind(data$mu0, data$mu1)
pi.hat <- data$pi1

fit <- grf_cate(
  data = data,
  modifiers = c("W1", "W2"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  mu.hat = mu.hat,
  pi.hat = pi.hat,
  cross_fit = FALSE
)
```

## Important Arguments

The GRF-specific arguments worth understanding are:

- `cross_fit` and `folds` for nuisance estimation
- `cross_validate` and `cv_control` for outer learner selection
- `ep_targeting_style` and `ep_r_targeting_basis` for the GRF-backed EP
  variants
- `treatment_level` and `control_level` when treatment labels are not the
  default `1` and `0`
- `grf_params` for GRF hyperparameters such as `num.trees` or
  `min.node.size`
- `tune` for the amount of GRF tuning: `"light"`, `"none"`, or `"all"`

In the GRF learners, `tune = "light"` tunes `sample.fraction`, `mtry`, and
`min.node.size` with a modest search budget. `tune = "none"` disables tuning,
and `tune = "all"` requests GRF's full tuning path.

## Important Constraints

- `grf_cate()` and `grf_crr()` currently support binary treatment only.
- `grf_crr()` requires a non-negative outcome.
- If you want the standard `hte_task()` plus `fit_*()` interface or
  continuous-treatment CATE, use the non-GRF workflow instead.

## Helper Constructors

When the wrappers are too narrow, these helpers expose the forest-backed
building blocks directly:

- `make_grf_base_learner()`
- `make_grf_nuisance_learners()`
- `make_grf_cate_learners()`
- `make_grf_crr_learners()`

Use these when you want manual cross-validation or a custom `sl3` learner
library rather than the default wrapper behavior.
