---
title: "Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`hte3` supports a three-step high-level workflow:

1. Create example or analysis data.
2. Build an `hte3_Task` with `hte_task()`.
3. Fit a model with `fit_cate()` or `fit_crr()`.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 120, seed = 1)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

model <- fit_cate(
  task,
  method = "dr",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(model, data))
summary(model)
```

## Cross-Fit vs Cross-Validate

Two arguments control different stages of the workflow:

- `cross_fit` in `hte_task()` controls nuisance estimation. When `TRUE`,
  nuisance learners such as the propensity score and outcome regression are
  cross-fitted.
- `cross_validate` in `fit_cate()` or `fit_crr()` controls model selection
  across candidate HTE learners.

These act at different stages of the pipeline. A common analysis uses
cross-fitting for nuisances and optionally uses HTE-level cross-validation to
select among candidate learners.

## Wrapper-Level Learner Portfolios

You can select among heterogeneous learner families directly from the wrapper
API by passing a vector to `method`.

```{r}
portfolio <- fit_cate(
  task,
  method = c("dr", "r", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

portfolio
head(predict(portfolio, data))
```

If `method = "ep"` and `cross_validate = TRUE`, `hte3` also cross-validates
across a portfolio of EP learners with different sieve basis sizes.

## CRR Workflow

The CRR wrapper uses the same pattern, but requires a non-negative outcome:

```{r}
crr_task <- hte_task(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y_binary",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

crr_portfolio <- fit_crr(
  crr_task,
  method = c("ipw", "t", "ep"),
  base_learner = Lrnr_mean$new(),
  cross_validate = TRUE,
  cv_control = list(V = 2)
)

head(predict(crr_portfolio, data))
```

See the other vignettes for CATE-specific, CRR-specific, low-level `sl3`
integration, and legacy reproducibility material.
