---
title: "Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Read this article if you want one minimal end-to-end example with the
high-level wrappers before choosing a more specific workflow.

```{r}
library(hte3)
library(sl3)

data <- hte3_example_data(n = 120, seed = 1)

task <- hte_task(
  data = data,
  modifiers = c("W1", "W2", "W3"),
  confounders = c("W1", "W2", "W3"),
  treatment = "A",
  outcome = "Y",
  propensity_learner = Lrnr_mean$new(),
  outcome_learner = Lrnr_mean$new(),
  mean_learner = Lrnr_mean$new(),
  cross_fit = FALSE
)

model <- fit_cate(
  task,
  method = "dr",
  base_learner = Lrnr_mean$new(),
  cross_validate = FALSE
)

head(predict(model, data))
summary(model)
```

## Cross-Fit vs Cross-Validate

Two arguments control different stages of the workflow:

- `cross_fit` in `hte_task()` controls nuisance estimation.
- `cross_validate` in `fit_cate()` or `fit_crr()` controls selection across
  candidate HTE learners.

These are separate decisions. A common analysis cross-fits nuisances and only
turns on wrapper-level cross-validation when comparing learner families.

## What to Read Next

- For CATE method choice, see [CATE Workflow](cate.html).
- For relative effects with non-negative outcomes, see [CRR Workflow](crr.html).
- For `grf_cate()` and `grf_crr()`, see [Causal Forests with GRF](grf.html).
- For low-level `sl3` control, see [Advanced sl3 Integration](advanced-sl3.html).
